Hydras \& Co.

### 2022-03-16 doc/part-hydras.tex
corrections chapter 4 (up to page 80)

@@

The ordinal \(\epsilon_0\) is the least ordinal number that satisfies 
the equation \(\alpha = \omega^\alpha\), where \(\omega\) is 
the least infinite ordinal. Thus, we can consider \(\epsilon_0\) as an
\emph{infinite} \(\omega\)-tower.
Nevertheless, 
any ordinal strictly less that \(\epsilon_0\) 
can be finitely represented by a unique  \emph{Cantor normal form}, 
that is, an expression  which is either  the ordinal \(0\) or 
a sum  \(\omega^{\alpha_1} \times n_1 + \omega^{\alpha_2} \times n_2 + 
  \dots + \omega^{\alpha_p} \times n_p\) where all the \(\alpha_i\) 
are ordinals in Cantor  normal form, \(\alpha_1 > \alpha_2 > \alpha_p\), 
and all the \(n_i\) are positive integers.

@@

\paragraph{Remark}
The name \texttt{T1} we gave to this data-type  is proper to this development and refers
to a hierarchy of ordinal notations. For instance, in Library \href{../theories/html/hydras.Gamma0.T2.html}{Gamma0.T2},  the following type is used to represent ordinals strictly less than \(\Gamma_0\),  in Veblen normal form (see also~\cite{schutte}).
\noindent

@@

\subsection{About Gaia}
\gaiasign Chapter~\vref{gaia-chapter} describes the present state of a
project of making compatible the libraries \HydrasLib and \gaia. At present, both libraries contain a version of the inductive type \texttt{T1}, with the same base name for the constructors.

@@

\paragraph*{\gaiasign}
  In \texttt{gaia.ssete9}, the $n$-th finite ordinal is written
  \texttt{$\backslash${F}\,$n$}. There is no coercion in \gaia from \texttt{nat} to finite ordinals.

@@

Note that \texttt{omega} is not an identifier, thus any tactic like \texttt{unfold omega} would fail.

\paragraph*{\gaiasign}
In \texttt{gaia.ssete9}, the ordinal $\omega$ is also named as \emph{constant}  \texttt{T1omega}.

@@

\paragraph*{\gaiasign}
In \texttt{gaia.ssete9} the identifier\texttt{phi0} is bound to a constant (not a notation).

@@

Let us consider again the ordinal $\alpha_0$ defined in section~\vref{alpha0-def}
If we ask \coq{} to print its  normal form, we get a hardly readable term of type \texttt{T1}.

@@

The following data type defines an abstract syntax for more readable ordinals terms in Cantor normal form:

@@

In Sect.~\vref{nf-gaia-compat}, we show that \gaia's \texttt{T1nf} is extensionnaly equivalent with \HydrasLib'  \texttt{nf}.

>>>>

The ordinal \(\epsilon_0\) is the least ordinal number that satisfies 
the equation \(\alpha = \omega^\alpha\), where \(\omega\) is 
the least infinite ordinal.
Thus, we can intuitively consider \(\epsilon_0\) as an
\emph{infinite} \(\omega\)-tower.

@@

Any ordinal strictly less that \(\epsilon_0\) 
can be finitely represented by a unique  \emph{Cantor normal form}, 
that is, an expression  which is either  the ordinal \(0\) or 
a sum  \(\omega^{\alpha_1} \times n_1 + \omega^{\alpha_2} \times n_2 + 
  \dots + \omega^{\alpha_p} \times n_p\) where all the \(\alpha_i\) 
are ordinals in Cantor  normal form, \(\alpha_1 > \alpha_2 > \alpha_p\), 
and all the \(n_i\) are positive integers.

@@

\paragraph{Remark}
The name \texttt{T1} we gave to this data-type  is proper to this development and refers
to a hierarchy of ordinal notations. For instance, in Library \href{../theories/html/hydras.Gamma0.T2.html}{Gamma0.T2},  the following type is used to represent ordinals strictly less than \(\Gamma_0\),  in Veblen normal form (see also~\cite{Gaia, schutte}).
\noindent

@@

\subsection{About Gaia}
\gaiasign Chapter~\vref{gaia-chapter} describes the present state of a
project of making compatible the libraries \HydrasLib and \gaia. At present, both libraries contain their own version of the inductive type \texttt{T1}, with the same base name for the constructors.

@@

\paragraph*{\gaiasign}
  In \texttt{gaia.ssete9}, the $n$-th finite ordinal is also written
  \texttt{$\backslash${F}\,$n$}. There is no coercion in \gaia from \texttt{nat} to finite ordinals.

@@

Note that \texttt{omega} is not an identifier in \HydrasLib, thus any tactic like \texttt{unfold omega} would fail.

\paragraph*{\gaiasign}
In \texttt{gaia.ssete9}, the ordinal $\omega$ is bound to the \emph{constant}  \texttt{T1omega}.

@@

\paragraph*{\gaiasign}
In \texttt{gaia.ssete9} the identifier \texttt{phi0} is bound to a plain constant (not a notation).

@@

Let us consider again the ordinal $\alpha_0$ defined in section~\vref{alpha0-def}. 
If we ask \coq{} to print its  normal form, we get a hardly readable term of type \texttt{T1}.

@@

The following data type defines a more readable abstract syntax for  ordinals terms in Cantor normal form:

@@

In Sect.~\vref{nf-gaia-compat}, we show that \gaia's \texttt{T1nf} is extensionally equivalent with \HydrasLib'  \texttt{nf}.
